---
title: "bimodality test essai"
output: html_notebook
---

We want to develop  new algorithm to automatically identify genes that are abnormally expressed in a subset of tumors : they are defined here as genes that have an unimodal profil of expression in normal tissues, and a bi- or multi-modal profil in tumors.

For C/T genes, we add the following criteria : the mean expression in normal tissue have to be below 1 FPKM-UQ.

# Intro
## Datas
Define the output directory
```{r}
setwd("~/Desktop/These_Marthe/1_Bioinfo/0_Scripts Propres")

pathRes="./res"
```

Load the data : 

```{r}
# SummarizedExperiment with expression data for TCGA breast samples, download from TCGA
data_brca <- readRDS(file = "./data/data_brca.rds")

#List of C/T genes, as the union of the 3 previously identified testis-specific genes
list_CT_genes <- read.table(file = "./data/list_CT_genes.txt", 
                            sep = "\t", header = TRUE)
```

##Functions
This function will calculate the gaussian kernel estimation of the density for each gene in NT and in Tumors samples, and will return a list of pic(s) and values for these pics. It will also calculate if this gene is abnormally expressed in tumors, according to our criteria (multi-modality in tumors and unimodality in NT)

```{r}
# X : vector of normalized expression
# Tum : index of tumors samples
# Tum : index of NT samples

F_derivative <- function(X, Tum, NT){
# Two vectors of expression : normal and tumors
  x_tum <- log2(1+ X[Tum])
  x_nt<- log2(1+ X[NT])
# Density for this C/T expression in NT and Tum
  d_tum <- density(x=x_tum, kernel="gaussian", bw = bw_fix)
  d_nt <- density(x=x_nt, kernel="gaussian", bw = bw_fix)


# dy/dx : first derivative
  first_tum <-diff(d_tum$y)/diff(d_tum$x)
  first_nt <-diff(d_nt$y)/diff(d_nt$x)

# to stock the sign changes of the derivative
  #For tumors : 
variations_tum <-c()
for(k in 2:length(first_tum)){
    if(sign(first_tum[k]) != sign(first_tum[k-1])){
         variations_tum <- c(variations_tum,k)
    }
}
  # For NT : 
variations_nt <-c()
for(k in 2:length(first_nt)){
    if(sign(first_nt[k]) != sign(first_nt[k-1])){
         variations_nt <- c(variations_nt,k)
    }
}


  
#Test : if the derivative sign change in tumor, but not in normal tissues => Significant
  if (length(variations_tum)>1 & length(variations_nt) == 1) {
    
    return(c(list(variations_tum), list(d_tum$x[variations_tum]),
           list(variations_nt), list(d_nt$x[variations_nt]),
            list("sign")))
    
  }else(  return(c(list(variations_tum), list(d_tum$x[variations_tum]),
           list(variations_nt), list(d_nt$x[variations_nt]),
            list("not sign"))))
   
}
```


This function will calculate the mean expression in tumors samples and in NT samples
```{r}
# X : vector of normalized expression
# Tum : index of tumors samples
# Tum : index of NT samples

F_mean <- function(X, Tum, NT){
    mean_Tum = mean(X[Tum], na.rm = TRUE)
    mean_NT = mean(X[NT], na.rm = TRUE)
    return(c(mean_NT, mean_Tum))
}
```



This function will be used in the "graphics" part

```{r}
## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=TRUE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm),
          max   = max     (xx[[col]], na.rm=na.rm),
          min   = min     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}
```

## Libraries
```{r message=FALSE, warning=FALSE}
library(SummarizedExperiment)
library(dplyr)
library(ggplot2)
```



#Iteration for C/T

Firest, we need to select the index for the C/T genes in data_brca. 

```{r}
id_gene_list<- which(rownames(assay(data_brca, 2)) %in% list_CT_genes$names == TRUE)

length(id_gene_list)
```

## Means

We need to have the mean for the second criteria of selection

```{r}
F_mean <- function(X, Tum, NT){
    mean_Tum = mean(X[Tum], na.rm = TRUE)
    mean_NT = mean(X[NT], na.rm = TRUE)
    return(c(mean_NT, mean_Tum))
}


```


```{r}
# For mean calculation : initialisation 
# -------------------------------------

# Tumors and normal tissues index
Tum <- which(colData(data_brca)$definition != "Solid Tissue Normal")
NT <- which(colData(data_brca)$definition == "Solid Tissue Normal")
Subtype <- colData(data_brca)$subtype_BRCA_Subtype_PAM50

# Subset of gene expression (FPKM-UQ) for CT genes only
gene_exp <- assay(data_brca, 2)[id_gene_list, ]*10^6


# calcul 
# ------

Table_mean <- apply(gene_exp, 1, F_mean, Tum = Tum, NT=NT ) %>% t %>% data.frame

colnames(Table_mean) <- c("mean_NT", "mean_Tum")
rownames(Table_mean) <- make.names(rowData(data_brca)[id_gene_list,"gene_id"], 
                                  unique= T)

Table_mean$Diff <- Table_mean$mean_Tum - Table_mean$mean_NT

head(Table_mean)
```


##Derivative


The bw parameter is important to play with the stringency of the screen : 

```{r}
# Initialisation
bw_fix = 0.7

Tum = which(colData(data_brca)$definition != "Solid Tissue Normal")
NT = which(colData(data_brca)$definition == "Solid Tissue Normal")

gene_exp <- assay(data_brca, 2)[id_gene_list, ]*10^6


# Apply the function ! 
res <- apply(gene_exp, 1, F_derivative, Tum = Tum, NT=NT )

# Organize the output into a list column, thanks to tibble
library(tibble)

res <- enframe(res) # converts a multi-level list to tibble with list cols

res$name <-  rownames(Table_mean)

head(res)
```

```{r}
library(purrr)
# To simplify the list column into a regular column, for character (sign / not sign)
sign_fun <- function(X){
  X[[5]]
}

label_sig <- res %>% transmute(name, output = map_chr(value, sign_fun))

head(label_sig)
```
```{r}
# To simplify the list column into a regular list : indices for pics
tum_fun <- function(X){
  X[[1]]
}

list_variation_tum <- res %>% transmute(name, output = map(value, tum_fun))
names(list_variation_tum$output) <- list_variation_tum$name
list_variation_tum <- list_variation_tum$output


head(list_variation_tum)
```
```{r}
# To simplify the list column into a regular list : values for pics
tum_fun <- function(X){
  X[[2]]
}

max_variation_tum <- res %>% transmute(name, output = map(value, tum_fun))
names(max_variation_tum$output) <- max_variation_tum$name
max_variation_tum <- max_variation_tum$output


head(max_variation_tum)
```
```{r}
# To simplify the list column into a regular list : indices for pics
nt_fun <- function(X){
  X[[3]]
}

list_variation_nt <- res %>% transmute(name, output = map(value, nt_fun))
names(list_variation_nt$output) <- list_variation_nt$name
list_variation_nt <- list_variation_nt$output


head(list_variation_nt)
```
```{r}
# To simplify the list column into a regular list : values for pics
nt_fun <- function(X){
  X[[4]]
}

max_variation_nt <- res %>% transmute(name, output = map(value, nt_fun))
names(max_variation_nt$output) <- max_variation_nt$name
max_variation_nt <- max_variation_nt$output


head(max_variation_nt)
```


```{r}

print(paste("We have identified", sum(label_sig$output=="sign"), 
            "multimodal TSPS bimodaux with bw =", bw_fix))


sort(label_sig$name[which(label_sig$output=="sign")])


```

2d criteria : we want to keep only the significatives C/T genes with a mean in NT < 1

```{r}
# How many ? 
which(Table_mean$mean_NT < 1 & label_sig$output =="sign") %>% length
```
```{r}
label_sig$mean <- "Notsign_OR_NotInf_1"
label_sig$mean[which(Table_mean$mean_NT < 1 & label_sig$output=="sign")] <- "sign_Inf1"


head(label_sig)
```


# Iteration for all genes


```{r}
# Initialisation
bw_fix = 0.7

Tum = which(colData(data_brca)$definition != "Solid Tissue Normal")
NT = which(colData(data_brca)$definition == "Solid Tissue Normal")

gene_exp <- assay(data_brca, 2)[, ]*10^6


# Apply the function ! 
res <- apply(gene_exp, 1, F_derivative, Tum = Tum, NT=NT )

# Organize the output into a list column, thanks to tibble
library(tibble)

res <- enframe(res) # converts a multi-level list to tibble with list cols

res$name <-  rownames(gene_exp)

head(res)
```
```{r}
# To simplify the list column into a regular column, for character (sign / not sign)
sign_fun <- function(X){
  X[[5]]
}

label_sig_all <- res %>% transmute(name, output = map_chr(value, sign_fun))

head(label_sig_all)
```
```{r}
# To simplify the list column into a regular list : indices for pics
tum_fun <- function(X){
  X[[1]]
}

list_variation_tum_all <- res %>% transmute(name, output = map(value, tum_fun))
names(list_variation_tum_all$output) <- list_variation_tum_all$name
list_variation_tum_all <- list_variation_tum_all$output


head(list_variation_tum_all)
```

```{r}
print(paste("on a identifié", sum(label_sig_all$output=="sign"), 
            "genes bimodaux avec ces paramètres, bw =", bw_fix))
```


```{r}
# To simplify the list column into a regular list : values for pics
tum_fun <- function(X){
  X[[2]]
}

max_variation_tum_all <- res %>% transmute(name, output = map(value, tum_fun))
names(max_variation_tum_all$output) <- max_variation_tum_all$name
max_variation_tum_all <- max_variation_tum_all$output


head(max_variation_tum_all)
```
```{r}
# To simplify the list column into a regular list : indices for pics
nt_fun <- function(X){
  X[[3]]
}

list_variation_nt_all <- res %>% transmute(name, output = map(value, nt_fun))
names(list_variation_nt_all$output) <- list_variation_nt_all$name
list_variation_nt_all <- list_variation_nt_all$output


head(list_variation_nt_all)
```
```{r}
# To simplify the list column into a regular list : values for pics
nt_fun <- function(X){
  X[[4]]
}

max_variation_nt_all <- res %>% transmute(name, output = map(value, nt_fun))
names(max_variation_nt_all$output) <- max_variation_nt_all$name
max_variation_nt_all <- max_variation_nt_all$output


head(max_variation_nt_all)
```



# Graphics


##  Density curves

To illustrate some specific exemples of gene expression levels distributions in normal tissues VS Tumors.
For Figure Supp1.

```{r}

# Graphical parameter
alpha = 1
bw_fix = 0.7

# Here put the name of your gene of interest
gene = "CHMP4B"

id_gene <- grep(gene, rownames(assay(data_brca, 2)))[1]

Gene_exp <- assay(data_brca, 2)[id_gene, ]*10^6

data=data.frame(exp=Gene_exp,
                label = colData(data_brca)$definition,
                Subt = Subtype)
data$label<- as.character(data$label)
data$label[which(data$label == "Primary solid Tumor")] = "Tum"
data$label[which(data$label == "Metastatic")] = "Tum"
data$label[which(data$label == "Solid Tissue Normal")] = "NT"


#variation de la dérivée
 d <- density(x=log2(1+data$exp[which(data$label != "NT")]), kernel="gaussian", bw = bw_fix)
# dy/dx first derivative
  first<-diff(d$y)/diff(d$x)
plot(first)
abline(h = 0, col = "red")


g1 = ggplot(data, aes(x = log2(1+Gene_exp), fill = label, color = label))+
 # geom_histogram(aes(y=..density..), alpha = alpha)+
  geom_density(bw = bw_fix, alpha = alpha)+
  scale_color_manual(values = c("orange", "grey10"))+
  scale_fill_manual(values = c("orange", "grey10"))+
  scale_x_continuous(limits = c(-2, 18))+
  theme_bw()+
  facet_grid(rows = vars(label))+

  labs(title= "",x="Expression values",cex=14)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12) ,
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x = element_text(face="bold"),
              axis.text.y=element_text(colour="black", size = 12))


g1


ggsave(paste(pathRes,"/",gene,"_", bw_fix, ".eps", sep=""), device=cairo_ps, g1)

```


## Heatmap

To illustrate the expression pattern of the most variables selected TS genes. Supposed to be coherent with PCA. Fig Supp1 ? 

```{r}
library(ComplexHeatmap)
library(circlize)

names_sign <- subset(label_sig, mean == "sign_Inf1")$name

id_row <- which(make.names(rowData(data_brca)$gene_id, unique = T) %in% names_sign)

# We take  the normalized expression value 
dat <- assay(data_brca, 2)[id_row,]*10^6
rownames(dat) <- make.names(rowData(data_brca)$gene_id, unique = T)[id_row]
head(dat)[, 1:5]
```

```{r}
# Select the most variable TS genes amoght the 139
top_var <- which(rownames(dat) %in% c(apply(dat, 1, var) %>% sort() %>% tail(., 40) %>% names()))

# we need a matrix for Heatmap function
mt <- as.matrix(log2(1+dat[top_var,]))


# Colors gradient
col_fun = colorRamp2(c( 0,   max(mt)), c("white", "blue"))


# columns annotations
ha = HeatmapAnnotation(Type = colData(data_brca)$definition,
                       Subtype = Subtype,
    col = list(Type = c("Metastatic" = "red", "Primary solid Tumor" = "black", 
                        "Solid Tissue Normal" = "white"),
               Subtype = c("Basal" = "hotpink4", "Her2" = "cyan4", "LumA" = "darkgoldenrod", "LumB" = "orange", "Normal" = "white")))

# EXpression Heatmap
ht_list = Heatmap(as.matrix(mt),
        top_annotation = ha,
        cluster_rows = TRUE,
        name = "Expression (log2)",
        col = col_fun,
        show_column_names = FALSE,
        row_title = "Top variable Selected Testis-Specific Genes",
        row_names_gp = gpar(fontsize = 7), 
       heatmap_legend_param = list(legend_direction = "horizontal"))

draw(ht_list, heatmap_legend_side = "bottom", 
    annotation_legend_side = "right")
```


## Heatmap 2

To illustrate the expression pattern of the most differentially ex^ressed between NT and Tumors selected TS genes. 

Fig Supp1 ? 


```{r}
Table_mean_sign <- subset(Table_mean, rownames(Table_mean) %in% names_sign)

```

```{r}
# Select the most variable TS genes amoght the 139
top_mean <- which(rownames(dat) %in% c(Table_mean_sign[ order(Table_mean_sign$Diff),] 
                                       %>% tail(., 15) %>% rownames))

# we need a matrix for Heatmap function
mt <- as.matrix(log2(1+dat[top_mean,]))


# Colors gradient
col_fun = colorRamp2(c( 0,   max(mt)), c("white", "blue"))


# columns annotations
ha = HeatmapAnnotation(Type = colData(data_brca)$definition,
                       Subtype = Subtype,
    col = list(Type = c("Metastatic" = "red", "Primary solid Tumor" = "black", 
                        "Solid Tissue Normal" = "white"),
               Subtype = c("Basal" = "hotpink4", "Her2" = "cyan4", "LumA" = "darkgoldenrod", "LumB" = "orange", "Normal" = "white")))

# EXpression Heatmap
ht_list = Heatmap(as.matrix(mt),
        top_annotation = ha,
        cluster_rows = TRUE,
        name = "Expression (log2)",
        col = col_fun,
        show_column_names = FALSE,
        row_title = "Top DeltaMean Selected Testis-Specific Genes",
        row_names_gp = gpar(fontsize = 7), 
       heatmap_legend_param = list(legend_direction = "horizontal"))

draw(ht_list, heatmap_legend_side = "bottom", 
    annotation_legend_side = "right")
```





```{r}
# Select the most variable TS genes amoght the 139
top_var <- which(rownames(dat) %in% c(apply(dat, 1, var) %>% sort() %>% tail(., 40) %>% names()))

# we need a matrix for Heatmap function
mt <- as.matrix(log2(1+dat[top_var,]))


# Colors gradient
col_fun = colorRamp2(c( 0,   max(mt)), c("white", "blue"))


# columns annotations
ha = HeatmapAnnotation(Type = colData(data_brca)$definition,
                       Subtype = Subtype,
    col = list(Type = c("Metastatic" = "red", "Primary solid Tumor" = "black", 
                        "Solid Tissue Normal" = "white"),
               Subtype = c("Basal" = "hotpink4", "Her2" = "cyan4", "LumA" = "darkgoldenrod", "LumB" = "orange", "Normal" = "white")))

# EXpression Heatmap
ht_list = Heatmap(as.matrix(mt),
        top_annotation = ha,
        cluster_rows = TRUE,
        name = "Expression (log2)",
        col = col_fun,
        show_column_names = FALSE,
        row_title = "Top variable Selected Testis-Specific Genes",
        row_names_gp = gpar(fontsize = 7), 
       heatmap_legend_param = list(legend_direction = "horizontal"))

draw(ht_list, heatmap_legend_side = "bottom", 
    annotation_legend_side = "right")
```


## Numer of modes and expression levels
### CT genes
```{r}
alpha = 1

Pic1_NT = c()
Pic1_tum = c()
Pic2_tum = c()
Pic2_NT = c()
Pic3_tum = c()
Pic3_NT = c()
Pic4_tum = c()
Pic4_NT = c()
Pic5_tum = c()
Pic5_NT = c()


for (i in 1:length( max_variation_nt )) {
   Pic1_NT <- c(Pic1_NT, max_variation_nt[[i]][1]) 
   Pic1_tum <- c(Pic1_tum, max_variation_tum[[i]][1])
   Pic2_tum <- c(Pic2_tum, max_variation_tum[[i]][3]) 
   Pic2_NT <- c(Pic2_NT, max_variation_nt[[i]][3]) 
   Pic3_tum <- c(Pic3_tum, max_variation_tum[[i]][5]) 
   Pic3_NT <- c(Pic3_NT, max_variation_nt[[i]][5]) 
   Pic4_tum <- c(Pic4_tum, max_variation_tum[[i]][7]) 
   Pic4_NT <- c(Pic4_NT, max_variation_nt[[i]][7]) 
      Pic5_tum <- c(Pic5_tum, max_variation_tum[[i]][9]) 
   Pic5_NT <- c(Pic5_NT, max_variation_nt[[i]][9]) 


}

data <- data.frame(pic = c(Pic1_NT, Pic2_NT,Pic3_NT,Pic4_NT, Pic5_NT,
                           Pic1_tum, Pic2_tum, Pic3_tum, Pic4_tum, Pic5_tum),
                   type = c(rep("NT", 5*length(Pic1_tum)),
                            rep("tum", 5* length(Pic1_tum))),
                   number = rep(c(rep("1", length(Pic1_tum)),
                                  rep("2", length(Pic1_tum)),
                                  rep("3", length(Pic1_tum)),
                                  rep("4", length(Pic1_tum)),
                                  rep("5", length(Pic1_tum))), 2),
                   label = rep(label_sig$output,  10))

tgc <- summarySE(data, measurevar="pic", groupvars=c("type","number", "label"))
for (i in 1:ncol(tgc)) {
  tgc[ which(is.infinite(tgc[,i])==T), i] <- NA

}

g3 = ggplot(data = tgc, 
       aes(y = pic, x = number, fill = type, color = type))+
  geom_errorbar(aes(ymin=min, ymax=max), width=.5, size = 1, linetype = 3) +
  geom_errorbar(aes(ymin=pic-sd, ymax=pic+sd), width=0.5, size = 1, linetype = 1) +

  geom_point(size = 5, shape = 18) + 
  #coord_flip() + 
  scale_color_manual(values = c("orange", "grey10"))+
  scale_fill_manual(values = c("blue", "grey10"))+
  facet_grid(rows = vars(label))+
  theme_bw()+

  labs(title= "",x="Relative maximum",cex=14)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12) ,
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x = element_text(face="bold"),
              axis.text.y=element_text(colour="black", size = 12))


g3
```


### All genes
```{r}
alpha = 1

Pic1_NT = c()
Pic1_tum = c()
Pic2_tum = c()
Pic2_NT = c()
Pic3_tum = c()
Pic3_NT = c()
Pic4_tum = c()
Pic4_NT = c()
Pic5_tum = c()
Pic5_NT = c()


for (i in 1:length( max_variation_nt_all )) {
   Pic1_NT <- c(Pic1_NT, max_variation_nt_all[[i]][1]) 
   Pic1_tum <- c(Pic1_tum, max_variation_tum_all[[i]][1])
   Pic2_tum <- c(Pic2_tum, max_variation_tum_all[[i]][3]) 
   Pic2_NT <- c(Pic2_NT, max_variation_nt_all[[i]][3]) 
   Pic3_tum <- c(Pic3_tum, max_variation_tum_all[[i]][5]) 
   Pic3_NT <- c(Pic3_NT, max_variation_nt_all[[i]][5]) 
   Pic4_tum <- c(Pic4_tum, max_variation_tum_all[[i]][7]) 
   Pic4_NT <- c(Pic4_NT, max_variation_nt_all[[i]][7]) 
      Pic5_tum <- c(Pic5_tum, max_variation_tum_all[[i]][9]) 
   Pic5_NT <- c(Pic5_NT, max_variation_nt_all[[i]][9]) 


}

data <- data.frame(pic = c(Pic1_NT, Pic2_NT,Pic3_NT,Pic4_NT, Pic5_NT,
                           Pic1_tum, Pic2_tum, Pic3_tum, Pic4_tum, Pic5_tum),
                   type = c(rep("NT", 5*length(Pic1_tum)),
                            rep("tum", 5* length(Pic1_tum))),
                   number = rep(c(rep("1", length(Pic1_tum)),
                                  rep("2", length(Pic1_tum)),
                                  rep("3", length(Pic1_tum)),
                                  rep("4", length(Pic1_tum)),
                                  rep("5", length(Pic1_tum))), 2),
                   label = rep(label_sig_all$output,  10))

tgc <- as.data.frame(summarySE(data, measurevar="pic", groupvars=c("type","number", "label")))

for (i in 1:ncol(tgc)) {
  tgc[ which(is.infinite(tgc[,i])==T), i] <- NA

}

g3 = ggplot(data = tgc, 
       aes(y = pic, x = number, fill = type, color = type))+
  geom_errorbar(aes(ymin=min, ymax=max), width=.5, size = 1, linetype = 3) +
  geom_errorbar(aes(ymin=pic-sd, ymax=pic+sd), width=0.5, size = 1, linetype = 1) +

  geom_point(size = 5, shape = 18) + 
  #coord_flip() + 
  scale_color_manual(values = c("orange", "grey10"))+
  scale_fill_manual(values = c("blue", "grey10"))+
  facet_grid(rows = vars(label))+
  theme_bw()+

  labs(title= "",x="Relative maximum",cex=14)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12) ,
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x = element_text(face="bold"),
              axis.text.y=element_text(colour="black", size = 12))


g3

```


```{r}
g3 = ggplot(data = tgc, 
       aes(y = pic, x = number, fill = type, color = type))+
  geom_errorbar(aes(ymin=min, ymax=max), width=.5, size = 1) +
  geom_point(size = 3) + 
  #coord_flip() + 
  scale_color_manual(values = c("orange", "grey10"))+
  scale_fill_manual(values = c("orange", "grey10"))+
  facet_grid(rows = vars(label))+
  theme_bw()+

  labs(title= "",x="Relative maximum",cex=14)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12) ,
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x = element_text(face="bold"),
              axis.text.y=element_text(colour="black", size = 12))


g3
```






# Write tables
```{r}

#POur les all
id_int = which(label_sig_all$output == "sign")

rowData(data_brca)[id_int, 1:3]

data <- data.frame(rowData(data_brca)[id_int, 1:3])
write.table(data, paste(pathRes, "Selected_All.txt", sep = "/"), quote = F, row.names = T, col.names = T, sep = "\t")


#pour les CT
id_int = which(label_sig_all$output == "sign" & is.element(label_sig_all$name,
                                                    list_CT_genes$names)==TRUE)
data <- data.frame(rowData(data_brca)[id_int, 1:3])
head(data)
write.table(data, paste(pathRes, "Selected_CT.txt", sep = "/"), quote = F, row.names = T, col.names = T, sep = "\t")

# Pour les CT avec critères de moyennes
list_Mean_all <- apply(assay(data_brca, 2)[, NT]*10^6, 1, function(X){mean(X, na.rm = T)})


id_int = which(list_Mean_all < 1 & label_sig_all$output == "sign" & is.element(label_sig_all$name,
                                                                        list_CT_genes$names)==TRUE)
data <- data.frame(rowData(data_brca)[id_int, 1:3])
head(data)
write.table(data, paste(pathRes, "Selected_CT_MeanInf1.txt", sep = "/"), quote = F, row.names = T, col.names = T, sep = "\t")


```








# Graphics : parameter bw

To highlight the importance of bw choice : we have tested different values for bw. The number of significant genes are keept here, and used to make the following representations.

```{r}
All = c( 978, 2430, 4612, 2871, 565, 389, 655,
         5422, 2994, 1433, 726)
CT = c(158, 312, 447, 472, 116, 76, 193,
       503, 325, 229, 126)
Data = data.frame(valeurs = c(All-CT, CT),
                  Param = rep(c(0.7, 0.5, 0.3, 0.1, 0.9, 0.99, 0.01,
                                0.2, 0.4, 0.6, 0.8), 2),
                  Type = c(rep("All", length(All)),
                           rep("CT", length(CT))))
Data$Param = as.factor(Data$Param)

ggplot(data=Data, aes(x=Param, y=valeurs, fill=Type)) +
  geom_bar(stat="identity", color = "black")+
  scale_fill_manual(values = c("grey80", "blue"))+
  scale_x_discrete( limits = c(0.01," ", 0.1," ", 0.2," ", 0.3," ", 0.4," ", 0.5," ", 0.6, " ",
                                0.7, " ",0.8, " ",0.9, " ",0.99))+
  theme_bw()+

  labs(title= "",x="Expression values",cex=14)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12) ,
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x = element_text(face="bold"),
              axis.text.y=element_text(colour="black", size = 12))
```

```{r}
nrow(data_brca)
nrow(list_CT_genes)
Data = data.frame(valeurs = c((rep(nrow(data_brca), 11)-All)/nrow(data_brca),
                              (rep(nrow(list_CT_genes),11)-CT)/nrow(list_CT_genes),
                              All/nrow(data_brca), CT/nrow(list_CT_genes)),
                  Select = c(rep("Unselected", 22),
                             rep("Selected", 22)),
                  Param = rep(c(0.7, 0.5, 0.3, 0.1, 0.9, 0.99, 0.01,
                                0.2, 0.4, 0.6, 0.8), 4),
                  Type = rep(c(rep("All", length(All)),
                           rep("CT", length(CT))), 2))
head(Data)
Data$Param = as.factor(Data$Param)
Data$valeurs = Data$valeurs*100


ggplot(data=Data, aes(x="", y=valeurs, fill=Select)) +
  geom_bar(stat="identity", color = "black")+ coord_polar("y", start=0)+
 facet_grid(rows = vars(Type), cols = vars(Param)) +
  scale_fill_manual(values = c("grey80", "white"))+
  theme_bw()+

  labs(title= "",x="Expression values",cex=14)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12) ,
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x=element_blank(),
              axis.text.y=element_text(colour="black", size = 12))
```



##Chi Test

In all case, we have significantly more Testis genes selected than "all genes" selected. This result confirm previous studies showing that Testis genes are promising markers in cancer.
```{r}

All_Unsel = rep(nrow(data_brca), 11)-All
All_Sel = All

CT_Unsel = rep(nrow(list_CT_genes),11)-CT
CT_Sel = CT

res_chi = NULL

for (i in 1:length(CT)) {
  data = matrix(c(All_Unsel[i],CT_Unsel[i], All_Sel[i], CT_Sel[i]), 
              ncol = 2) 
  res_chi = c(res_chi, 
              list(chisq.test(x=data)))}
res_chi

```







