---
title: "R Notebook"
output: html_notebook
---

We want to develop  new algorithm to automatically identify genes that are abnormally expressed in a subset of tumors : they are defined here as genes that have an unimodal profil of expression in normal tissues, and a bi- or multi-modal profil in tumors.

For C/T genes, we add the following criteria : the mean expression in normal tissue have to be below 1 FPKM-UQ.

# Intro
## Datas
Define the output directory
```{r}
setwd("~/Desktop/These_Marthe/1_Bioinfo/0_Scripts Propres")

pathRes="./res"
```

Load the data : 

```{r}
# SummarizedExperiment with expression data for TCGA breast samples, download from TCGA
data_brca <- readRDS(file = "./data/TCGA_Breast_data_FPKMUQv4.rds")

#List of C/T genes, as the union of the 3 previously identified testis-specific genes
list_CT_genes <- read.table(file = "./data/list_CT_genes.txt", 
                            sep = "\t", header = TRUE)
```

##Functions
This function will calculate the gaussian kernel estimation of the density for each gene in NT and in Tumors samples, and will return a list of pic(s) and values for these pics. It will also calculate if this gene is abnormally expressed in tumors, according to our criteria (multi-modality in tumors and unimodality in NT)

```{r}
# X : vector of normalized expression
# Tum : index of tumors samples
# Tum : index of NT samples

F_derivative <- function(X, Tum, NT){
# Two vectors of expression : normal and tumors
  x_tum <- log2(1+ X[Tum])
  x_nt<- log2(1+ X[NT])
# Density for this C/T expression in NT and Tum
  d_tum <- density(x=x_tum, kernel="gaussian", bw = bw_fix)
  d_nt <- density(x=x_nt, kernel="gaussian", bw = bw_fix)


# dy/dx : first derivative
  first_tum <-diff(d_tum$y)/diff(d_tum$x)
  first_nt <-diff(d_nt$y)/diff(d_nt$x)

# to stock the sign changes of the derivative
  #For tumors : 
variations_tum <-c()
for(k in 2:length(first_tum)){
    if(sign(first_tum[k]) != sign(first_tum[k-1])){
         variations_tum <- c(variations_tum,k)
    }
}
  # For NT : 
variations_nt <-c()
for(k in 2:length(first_nt)){
    if(sign(first_nt[k]) != sign(first_nt[k-1])){
         variations_nt <- c(variations_nt,k)
    }
}


  
#Test : if the derivative sign change in tumor, but not in normal tissues => Significant
  if (length(variations_tum)>1 & length(variations_nt) == 1) {
    
    return(c(list(variations_tum), list(d_tum$x[variations_tum]),
           list(variations_nt), list(d_nt$x[variations_nt]),
            list("sign")))
    
  }else(  return(c(list(variations_tum), list(d_tum$x[variations_tum]),
           list(variations_nt), list(d_nt$x[variations_nt]),
            list("not sign"))))
   
}
```


This function will calculate the mean expression in tumors samples and in NT samples
```{r}
# X : vector of normalized expression
# Tum : index of tumors samples
# Tum : index of NT samples

F_mean <- function(X, Tum, NT){
    mean_Tum = mean(X[Tum], na.rm = TRUE)
    mean_NT = mean(X[NT], na.rm = TRUE)
    return(c(mean_NT, mean_Tum))
}
```



This function will be used in the "graphics" part

```{r}
## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=TRUE,
                      conf.interval=.95, .drop=TRUE) {
    library(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm),
          max   = max     (xx[[col]], na.rm=na.rm),
          min   = min     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}
```

## Libraries
```{r message=FALSE, warning=FALSE}
library(SummarizedExperiment)
library(dplyr)
library(ggplot2)
```



#Iteration for tested genes

```{r}
list_genes_test <- c("ESR1", "PGR", "ERBB2", 
                     "ACTB","GAPDH", "TUBA1A")
```


First, we need to select the index for the C/T genes in data_brca. 

```{r}
id_gene_list<- which(rowData(data_brca)$external_gene_name  %in% list_genes_test == TRUE)

length(id_gene_list)
```

## Means

We need to have the mean for the second criteria of selection

```{r}
F_mean <- function(X, Tum, NT){
    mean_Tum = mean(X[Tum], na.rm = TRUE)
    mean_NT = mean(X[NT], na.rm = TRUE)
    return(c(mean_NT, mean_Tum))
}


```


```{r}
# For mean calculation : initialisation 
# -------------------------------------

# Tumors and normal tissues index
Tum <- which(colData(data_brca)$definition != "Solid Tissue Normal")
NT <- which(colData(data_brca)$definition == "Solid Tissue Normal")
Subtype <- colData(data_brca)$subtype_BRCA_Subtype_PAM50

# Subset of gene expression (FPKM-UQ) for CT genes only
gene_exp <- assay(data_brca)[id_gene_list, ]*10^6


# calcul 
# ------

Table_mean <- apply(gene_exp, 1, F_mean, Tum = Tum, NT=NT ) %>% t %>% data.frame

colnames(Table_mean) <- c("mean_NT", "mean_Tum")
rownames(Table_mean) <- make.names(rowData(data_brca)[id_gene_list,"external_gene_name"], 
                                  unique= T)

Table_mean$Diff <- Table_mean$mean_Tum - Table_mean$mean_NT

head(Table_mean)
```


##Derivative


The bw parameter is important to play with the stringency of the screen : 

```{r}
list_label_sig <- list()

for (k in seq(0.1,1, 0.1)) {
  
# Initialisation
bw_fix = k

Tum = which(colData(data_brca)$definition != "Solid Tissue Normal")
NT = which(colData(data_brca)$definition == "Solid Tissue Normal")

gene_exp <- assay(data_brca)[id_gene_list, ]*10^6


# Apply the function ! 
res <- apply(gene_exp, 1, F_derivative, Tum = Tum, NT=NT )

# Organize the output into a list column, thanks to tibble
library(tibble)

res <- enframe(res) # converts a multi-level list to tibble with list cols

res$name <-  rownames(Table_mean)


library(purrr)
# To simplify the list column into a regular column, for character (sign / not sign)
sign_fun <- function(X){
  X[[5]]
}

label_sig <- res %>% transmute(name, output = map_chr(value, sign_fun))

head(label_sig)


list_label_sig <- c(list_label_sig, label_sig)

}

list_label_sig


```

# Graphics


##  Density curves

To illustrate some specific exemples of gene expression levels distributions in normal tissues VS Tumors.
For Figure Supp1.

```{r}
for (k in seq(0.1,1, 0.1)) {

# Graphical parameter
alpha = 1
bw_fix = k

# Here put the name of your gene of interest
gene = "ERBB2"

id_gene <-which(rowData(data_brca)$external_gene_name == gene)

Gene_exp <- assay(data_brca)[id_gene, ]*10^6

data=data.frame(exp=Gene_exp,
                label = colData(data_brca)$definition)
data$label<- as.character(data$label)
data$label[which(data$label == "Primary solid Tumor")] = "Tum"
data$label[which(data$label == "Metastatic")] = "Tum"
data$label[which(data$label == "Solid Tissue Normal")] = "NT"


#variation de la d?riv?e
 d <- density(x=log2(1+data$exp[which(data$label != "NT")]), kernel="gaussian", bw = bw_fix)
# dy/dx first derivative
  first<-diff(d$y)/diff(d$x)
plot(first)
abline(h = 0, col = "red")


g1 = ggplot(data, aes(x = log2(1+Gene_exp), fill = label, color = label))+
 # geom_histogram(aes(y=..density..), alpha = alpha)+
  geom_density(bw = bw_fix, alpha = alpha)+
  scale_color_manual(values = c("grey80", "grey10"))+
  scale_fill_manual(values = c("grey80", "grey10"))+
  scale_x_continuous(limits = c(30, 50))+
  theme_bw()+
  facet_grid(rows = vars(label))+

  labs(title= "",x="Expression values",cex=14)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12) ,
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x = element_text(face="bold"),
              axis.text.y=element_text(colour="black", size = 12))


g1


ggsave(paste(pathRes,"/",gene,"_", bw_fix, ".eps", sep=""), device=cairo_ps, g1)

}

```
```{r}
 ggplot(data, aes(x = log2(1+Gene_exp), fill = label, color = label))+
  geom_histogram(aes(y =..density..), alpha = alpha, binwidth = 0.01 )+
  scale_color_manual(values = c("grey80", "grey10"))+
  scale_fill_manual(values = c("grey80", "grey10"))+
  scale_x_continuous(limits = c(30, 50))+
  theme_bw()+
  facet_grid(rows = vars(label))+

  labs(title= "",x="Expression values",cex=14)+
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12) ,
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x = element_text(face="bold"),
              axis.text.y=element_text(colour="black", size = 12))
```

```{r}
# To simplify the list column into a regular list : indices for pics
tum_fun <- function(X){
  X[[1]]
}

list_variation_tum <- res %>% transmute(name, output = map(value, tum_fun))
names(list_variation_tum$output) <- list_variation_tum$name
list_variation_tum <- list_variation_tum$output


head(list_variation_tum)
```
```{r}
# To simplify the list column into a regular list : values for pics
tum_fun <- function(X){
  X[[2]]
}

max_variation_tum <- res %>% transmute(name, output = map(value, tum_fun))
names(max_variation_tum$output) <- max_variation_tum$name
max_variation_tum <- max_variation_tum$output


head(max_variation_tum)
```
```{r}
# To simplify the list column into a regular list : indices for pics
nt_fun <- function(X){
  X[[3]]
}

list_variation_nt <- res %>% transmute(name, output = map(value, nt_fun))
names(list_variation_nt$output) <- list_variation_nt$name
list_variation_nt <- list_variation_nt$output


head(list_variation_nt)
```
```{r}
# To simplify the list column into a regular list : values for pics
nt_fun <- function(X){
  X[[4]]
}

max_variation_nt <- res %>% transmute(name, output = map(value, nt_fun))
names(max_variation_nt$output) <- max_variation_nt$name
max_variation_nt <- max_variation_nt$output


head(max_variation_nt)
```


```{r}

print(paste("We have identified", sum(label_sig$output=="sign"), 
            "multimodal genes with bw =", bw_fix))


sort(label_sig$name[which(label_sig$output=="sign")])


```

2d criteria : we want to keep only the significatives C/T genes with a mean in NT < 1

```{r}
# How many ? 
which(Table_mean$mean_NT < 1 & label_sig$output =="sign") %>% length
```
```{r}
label_sig$mean <- "Notsign_OR_NotInf_1"
label_sig$mean[which(Table_mean$mean_NT < 1 & label_sig$output=="sign")] <- "sign_Inf1"


head(label_sig)
```

