---
title: "R Notebook"
output: html_notebook
---

#Introduction

Objective : Assess the Testis-Specific pattern of expression for the selected TS genes.


Data : RNA-Seq from differents human tissues are downloaded from GTEx website (TPM).
In this dataframe, I have extracted data for annotated human tissues, and selected CT genes. 

##Librairies

```{r}
library(stringi)
library(dplyr)

library(ggplot2)

library(circlize)
library(ComplexHeatmap)
```

## Data

```{r}
setwd("~/Desktop/These_Marthe/1_Bioinfo/0_Scripts Propres")

dat <- read.table('./data/GTEx_SelectedTS.txt', header =TRUE, sep ='\t')
colnames(dat) <- stri_replace_all(colnames(dat),regex =  "\\.", replacement = "-")

selected <- read.table("./data/Selected_CT_MeanInf1.txt", sep = "\t", dec = ",", header = T)

anno_dat <- read.table("./data/GTEx_anno_SelectedTS.txt", 
                       header = TRUE, sep = '\t', na.strings = "")
anno_dat <- anno_dat[which( anno_dat$SAMPID %in% colnames(dat) ), ]

```



To check if annotation and data are in consistent order : 
```{r}
c(as.character(anno_dat$SAMPID) == as.character(colnames(dat))) %>% sum == ncol(dat)

```

# Boxplot 

Reprganize data in ggplot style : 

```{r}
dt_graph <- data.frame(stack(dat))

dt_graph$SMTS <- rep(anno_dat$SMTS, each = nrow(dat))
dt_graph$SMTSD <- rep(anno_dat$SMTSD, each = nrow(dat))

dt_graph$names <- rep(rownames(dat), times = ncol(dat))
head(dt_graph)
```

Compute median per gene and per tissue with aggregate function : 

```{r}
dt <- aggregate(values ~ SMTS + names, data = dt_graph, FUN= "median" )
dt$sd <- aggregate(values ~ SMTS + names, data = dt_graph, FUN= "sd" )$values
dt$sd_min <- dt$values - dt$sd 
dt$sd_pos <- dt$values + dt$sd 

head(dt)
```

Boxplot : median expression, log10 scale
```{r}
ggplot(data = dt, aes(x = SMTS, y = values, color = SMTS)) +
  geom_boxplot(outlier.colour = "white")+
 # geom_errorbar(data=dt, mapping=aes(x=SMTS, ymin=sd_min, ymax=sd_pos), width=0.2, size=1) +
  geom_jitter(shape = 19)+
  
  scale_color_manual(values = c(rep("black", 23), "red4", rep("black", 3)))+
  scale_y_log10()+
  labs(title  =  "139 Selected C/T genes:GTEX",  x="Organs",y="TPM")+ 
  
  theme_classic()+                                                     
  theme(plot.title = element_text(size = 12,hjust=0.5),   
              text=element_text(),
              axis.title = element_text( size=10),       
              axis.text.x=element_text(angle=45, hjust=1,colour="black", size = 8),
              axis.text.y=element_text(colour="black", size = 8),
              legend.position = "none")
```









# heatmap
```{r}
#Compute the median per TS gene and per tissue
dh <- apply(dat, 1, function(X){by(X, INDICES = anno_dat$SMTS, median)}) %>% t

#Remove 0 values
dh <-dh[-which(rowSums(dh)==0),]


# Scale the matrix (variable = genes, by row)
mt <- dh[,-1] %>% apply(., 1, scale) %>% t

#Annotations
rownames(mt) <- rownames(dh)
colnames(mt) <- colnames(dh)[-1]
ha = HeatmapAnnotation(Type = colnames(dh)[-1],
     gp = gpar(col = "black"))


# Colors gradient for expression
col_fun = colorRamp2(c( min(mt,na.rm = T),0,  max(mt, na.rm = T)), c("blue", "white", "red"))



# EXpression Heatmap
ht_list = Heatmap(as.matrix(mt),
        top_annotation = ha,
        name = "Expression (scale)",
        col = col_fun,
        rect_gp = gpar(col = "black", lwd = 0.1),
        show_column_names = TRUE,
        row_title = "Selected TS Genes",
        row_names_gp = gpar(fontsize = 4), 
       heatmap_legend_param = list(legend_direction = "horizontal"))

draw(ht_list, heatmap_legend_side = "bottom", 
    annotation_legend_side = "right")

```



