---
title: "R Notebook"
output: html_notebook
---

# Intro

We want to clusterize tumors according to the 139 selected testis gene expression. To do so, we will do a principal component analysis, using as supplementary variables the ColData. 

## Data
```{r}
setwd("~/Desktop/These_Marthe/1_Bioinfo/0_Scripts Propres")

pathRes="./res"
```



```{r}
Selected_CT <- read.delim("./data/Selected_CT_MeanInf1.txt", 
                            sep = "\t", header = TRUE)

# SummarizedExperiment with expression data for TCGA breast samples, download from TCGA
data_brca <- readRDS(file = "./data/data_brca.rds")

```


## Libraries

```{r message=FALSE, warning=FALSE}
library(FactoMineR)
library(factoextra)
library(dplyr)
library(tidyverse)
library(gridExtra)

library(SummarizedExperiment)

library(ComplexHeatmap)
library(circlize)

```


# I. PCA

Select the data : 139 Testis specific gene expression

```{r}
id <- which( is.element(rownames(data_brca), rownames(Selected_CT))==TRUE)
length(id)

data.active <- data.frame(t(assay(data_brca[id,], 2))*10^6)

# Log scale the data
data.active <- data.active %>% (function(X){log2(X+1)})

head(data.active[, 1:5])
```


Select the supplementary variables (they will not contribute to the PCs, but they will be correlated with them)

```{r}
data.supp <- colData(data_brca) %>% as.data.frame()

head(data.supp)
```
Select quantitative and qualitative supplementary variables : 
```{r}
id_quali <- which(summary(data.supp)[2, ]=="Class :character  ")

id_quanti <- setdiff(1:ncol(data.supp), id_quali)

#Transform character vector to factors : 
data.supp[, id_quali] <- lapply(data.supp[,id_quali], factor )
head(data.supp)
```

```{r}
# Add Normal Tissue information (instead of NA) to the subtype variable : 
data.supp$subtype_BRCA_Subtype_PAM50 = as.character(data.supp$subtype_BRCA_Subtype_PAM50)
data.supp$subtype_BRCA_Subtype_PAM50[which(data.supp$shortLetterCode=="NT")] = "NT"
data.supp$subtype_BRCA_Subtype_PAM50 = as.factor(data.supp$subtype_BRCA_Subtype_PAM50)

```

Set the data to the PCA & run it: 
```{r}
#Remove quantitative supplementary variables : 
data.tot <- cbind(data.active, data.supp[,-id_quanti])


res.pca.norm <- PCA(data.tot[, ], scale.unit = F, ncp = 50, graph = TRUE,
               quali.sup = c(1+ncol(data.active)):ncol(data.tot) )
```

Variance explained by the 20 first components : 
```{r}
fviz_eig(res.pca.norm, addlabels = TRUE, ylim = c(0, 50),  barfill = "grey", barcolor = "black", ncp = 20)+
  labs(title= "Variance Explained ", 
       y="Percentage of explained variances (%)",x="Dimension",cex=10)+
  theme_classic()+ 
 theme(plot.title = element_text(size = 14, face = "bold",hjust = 0.5),
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(angle=45, hjust=1,colour="black", size = 11),
              axis.text.y=element_text(colour="black", size = 11))
```


Variable contribution to the 2 first components : 
```{r}
fviz_contrib(res.pca.norm, choice = "var", axes = 1:2, fill = "grey",
  color = "black", top = 11)+
  labs(title= "Contribution of variables to Dim 1-2 ", 
       y="Constributions (%)",x="",cex=10)+
  theme_classic()+ 
 theme(plot.title = element_text(size = 14, face = "bold",hjust = 0.5),
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(angle=45, hjust=1,colour="black", size = 11),
              axis.text.y=element_text(colour="black", size = 11))
```

```{r}
# We will represent on the graph the Top10 variable contributing to the PC1 and PC2
# To divide by the number considered component, here 2:
div2 = function(X){X/2}

# TO pondere each component by the % of variability explained : 
poids <- res.pca.norm$eig[1:2, 2] / 100
pondere <- function(X){X*poids}

# Top 10 (same than with the embedied graphical function fviz_contrib)
name <- list(name = res.pca.norm$var$contrib[, 1:2] %>% pondere %>% rowSums %>% div2 %>% sort(.,decreasing = T) %>%  head(.,10) %>% names)

fviz_pca_biplot (res.pca.norm, axes = c(1, 2),
                 col.var = "black",
              col.ind = data.tot$subtype_BRCA_Subtype_PAM50,
              select.var = name, repel = TRUE,
                
              geom.ind  = c("point"), geom.var = c("arrow", "text"),
              pointsize = "cos2",  #palette = "jco", 
              pointshape = 21
             ) +
 scale_color_manual(values=c("hotpink4","darkcyan","goldenrod3", "orange3", "grey50", "white"))+
  theme_classic()
```

## UMAP 
Option for the representation of PCA (but we can't have the variable on the same graph)

```{r}
library(umap)

# Create Seurat object
custome.config <- umap.defaults
custome.config$n_neighbors = 250
custome.config$n_epochs = 500
custome.config$n_components = 2

dataUMAP <- umap(data.active,
                 custome.config)

dataGG <- data.frame(dataUMAP$layout)
ggplot(data = dataGG,
       aes ( x = X1,y =X2,
             color = data.tot$subtype_BRCA_Subtype_PAM50))+
  geom_point(size = 3, shape = 1)+
 scale_color_manual(values=c("hotpink4","darkcyan","goldenrod3", "orange3", "grey50", "grey70"))+
  theme_classic()+ 
 theme(plot.title = element_text(size = 14, face = "bold",hjust = 0.5),
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(angle=45, hjust=1,colour="black", size = 11),
              axis.text.y=element_text(colour="black", size = 11))
```

# II. Boxplot 

This part allow you to automatically draw boxplot of a given gene (here, HORMAD1 and CT83 / CXorf61) according to the tumor subtype. 

```{r}
A = "MAGEA8"
id <- grep(A, rownames(data_brca))

data <- data.frame(Exp = t(assay(data_brca[id,], 2))*10^6, 
                         Subtype = colData(data_brca)$subtype_BRCA_Subtype_PAM50,
                         subtype_pathologic_stage = colData(data_brca)$subtype_pathologic_stage, 
                          shortLetterCode = colData(data_brca)$shortLetterCode)
colnames(data)[1] = "Exp"

#Nommer les NT
data$Subtype = as.character(data$Subtype)
data$Subtype[which(colData(data_brca)$shortLetterCode == "NT")] = "NT"
data$Subtype = as.factor(data$Subtype)
unlist(summary(aov(formula = Exp~Subtype, data = data)))[9]

g1= ggplot(data = data, aes(x = Subtype, y = Exp+1, fill = Subtype, color = Subtype))+
  geom_boxplot(notch = TRUE, size = 0.5, outlier.colour = "white", color = "black")+
  geom_jitter( width = 0.1, shape = 1)+
  coord_trans(y="log2")+
  scale_x_discrete(limits= c("NT", "LumA", "LumB", "Her2", "Basal", "Normal") )+
  scale_fill_manual(values=c("hotpink4","darkcyan","goldenrod3", "orange3", "grey50", "white"))+
  scale_color_manual(values=c("hotpink3","cyan3","goldenrod2", "orange1", "grey70", "grey80"))+
  labs(title=paste(A), x="Subtype",y="FPKM-UQ")+                         
  theme_classic()+                                                     # thème blanc
  theme(plot.title = element_text(size = 12, face = "bold",hjust=0.5),    #titre en gras, centré
              text=element_text(),
              axis.title = element_text(face="bold", size=10),        #titre des axes en gras
              axis.text.x=element_text(angle=45, hjust=1,colour="black", size = 8),
              axis.text.y=element_text(colour="black", size = 8),
              legend.position = "none")
g1
```

```{r}
A = "CXorf61"
id <- grep(A, rownames(data_brca))

data <- data.frame(Exp = t(assay(data_brca[id,], 2))*10^6, 
                         Subtype = colData(data_brca)$subtype_BRCA_Subtype_PAM50,
                         subtype_pathologic_stage = colData(data_brca)$subtype_pathologic_stage, 
                          shortLetterCode = colData(data_brca)$shortLetterCode)
colnames(data)[1] = "Exp"

#Nommer les NT
data$Subtype = as.character(data$Subtype)
data$Subtype[which(colData(data_brca)$shortLetterCode == "NT")] = "NT"
data$Subtype = as.factor(data$Subtype)

unlist(summary(aov(formula = Exp~Subtype, data = data)))[9]

g2= ggplot(data = data, aes(x = Subtype, y = Exp+1, fill = Subtype, color = Subtype))+
  geom_boxplot(notch = TRUE, size = 0.5, outlier.colour = "white", color = "black")+
  geom_jitter( width = 0.1, shape = 1)+
  coord_trans(y="log2")+
  scale_x_discrete(limits= c("NT", "LumA", "LumB", "Her2", "Basal", "Normal") )+
  scale_fill_manual(values=c("hotpink4","darkcyan","goldenrod3", "orange3", "grey50", "white"))+
  scale_color_manual(values=c("hotpink3","cyan3","goldenrod2", "orange1", "grey70", "grey80"))+
  labs(title=paste(A), x="Subtype",y="FPKM-UQ")+                         
  theme_classic()+                                                     # thème blanc
  theme(plot.title = element_text(size = 12, face = "bold",hjust=0.5),    #titre en gras, centré
              text=element_text(),
              axis.title = element_text(face="bold", size=10),        #titre des axes en gras
              axis.text.x=element_text(angle=45, hjust=1,colour="black", size = 8),
              axis.text.y=element_text(colour="black", size = 8),
              legend.position = "none")
g2
```
```{r}
grid.arrange(g1, g2, ncol = 2)
```

# III. Biplot

This part will draw biplot of expression data of HORMAD1 vs. CT83 : 
1. in the all dataset
2. In the basal tumor subtype, afetr characterization of the ON / OFF statut of both genes. A testis-gene is considered as "ON" if the expression level in a given tumor tissue is above mean+3sd the expression level in normal tissues.

```{r}
id_gene <- c(grep("HORMAD1",rownames(data_brca) ),
             grep("CXorf61",rownames(data_brca) ))
data <- assay(data_brca, 2)[id_gene, ] %>% t %>% data.frame
data <- data*10^6
colnames(data) <- c("Hormad1", "CT83")
data$type <- data.tot$subtype_BRCA_Subtype_PAM50

# On 
data <- data[order(rownames(data)),]

library(ggplot2)

ggplot(data = data, aes(x = log2(1+Hormad1), y = log2(1+CT83), fill = type, color = type ))+
  geom_point(size = 3, shape = 21, color = "black", na.rm = T) + 
  geom_rug(outside = TRUE)+
  coord_cartesian(clip = "off")+

  scale_fill_manual(values=c("hotpink4","darkcyan","goldenrod3", "orange3", "grey50", "white"))+
  scale_color_manual(values=c("hotpink4","darkcyan","goldenrod3", "orange3", "grey50", "white"))+

  theme_classic()+ 
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12) ,
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(angle=45, hjust=1,colour="black", size = 10),
              axis.text.y=element_text(colour="black", size = 10),
      legend.position="right")
```

Compute the ON/ OFF statut

```{r}
id_NT <- which(data$type == "NT")

Hormad_CT83_ON <- which(data$Hormad1 > c(mean(data$Hormad1[id_NT])+3*sd(data$Hormad1[id_NT])) & 
                     data$CT83 > c(mean(data$CT83[id_NT])+3*sd(data$CT83[id_NT]))&
                     data$type == "Basal")

Hormad_ON <- which(data$Hormad1 > c(mean(data$Hormad1[id_NT])+3*sd(data$Hormad1[id_NT])) & 
                     data$CT83 <= c(mean(data$CT83[id_NT])+3*sd(data$CT83[id_NT]))&
                     data$type == "Basal")

CT83_ON <- which(data$Hormad1 <= c(mean(data$Hormad1[id_NT])+3*sd(data$Hormad1[id_NT])) & 
                     data$CT83 > c(mean(data$CT83[id_NT])+3*sd(data$CT83[id_NT]))&
                     data$type == "Basal")

Both_OFF <- which(data$Hormad1 <= c(mean(data$Hormad1[id_NT])+3*sd(data$Hormad1[id_NT])) & 
                     data$CT83 <= c(mean(data$CT83[id_NT])+3*sd(data$CT83[id_NT]))&
                     data$type == "Basal")

Infos <- data.frame(Index = c(Hormad_CT83_ON, Hormad_ON , CT83_ON, Both_OFF),
                    Type = c(rep("Hormad_CT83_ON", length(Hormad_CT83_ON)),
                             rep("Hormad_ON", length(Hormad_ON)),
                             rep("CT83_ON", length(CT83_ON)),
                             rep("Both_OFF", length(Both_OFF))))



table(Infos$Type)
```


Select basal tumors : 
```{r}
ggplot(data = data[Infos$Index,], aes(x = log2(1+Hormad1), y = log2(1+CT83),
                                      fill = type, color = type ,
                                      shape = Infos$Type ))+
  geom_point(size = 3, color  = "black") + 
  geom_rug(outside = TRUE)+
  coord_cartesian(clip = "off")+

  scale_fill_manual(values=c("hotpink4","darkcyan","goldenrod3", "orange3", "grey50", "white"))+
  scale_color_manual(values=c("hotpink4","darkcyan","goldenrod3", "orange3", "grey50", "white"))+
  scale_shape_manual(values = c(21, 22, 23, 24))+

  theme_classic()+ 
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 12) ,
              text=element_text(),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(angle=45, hjust=1,colour="black", size = 10),
              axis.text.y=element_text(colour="black", size = 10),
      legend.position="right")
```



